---
title: "Notebook"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Loading needed libraries

```{r, message = FALSE}
library("DESeq2")
library("IHW")
library('ggplot2')
library('regionReport')
library("RColorBrewer") 
library("pheatmap")
library("vsn")
```

### Tidying up the data

```{r, results='hide', message=FALSE, warning=FALSE}
counts = read.delim("~/MScBioinformatics/Period4/GA/stunning-garbanzo/results/transcriptomics/Counts/htseq_all_sample_count.tsv", header=F, row.names=1)
numc <- sapply(counts, is.factor)
counts[numc] <- lapply(counts[numc], function(x) as.numeric(as.character(x)))
counts <- na.omit(counts)
colnames(counts) = c("BH1", "SR1", "BH2", "SR2", "BH3", "SR3")
my.design <- data.frame(
  row.names = colnames( counts ),
  condition = c( "BH", "Serum", "BH", "Serum","BH", "Serum"),
  libType = c( "paired-end", "paired-end","paired-end","paired-end","paired-end","paired-end" )
)
conds <- factor(my.design$condition)
dds <- DESeqDataSetFromMatrix(countData = counts, colData = my.design, design = ~ condition)
featureData <- data.frame(gene=rownames(counts))
mcols(dds) <- DataFrame(mcols(dds), featureData)
```


### Pre-Filtering for reads with at least 10 counts
```{r, messsage = FALSE}
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
```
### Refactor control and BH
```{r, message = FALSE}
dds$condition <- relevel(dds$condition, ref = "BH")
```

### Differential Expression analysis
```{r, message = FALSE}
dds <- DESeq(dds)
res <- results(dds, filterFun = ihw)
resOrdered <- res[order(res$pvalue),]
```

### Top 50 results
```{r}
n = 50
topResults <- rbind( resOrdered[ abs(resOrdered[,'log2FoldChange']) > 2, ][1:n,], 
                     resOrdered[ abs(resOrdered[,'log2FoldChange']) < 0.5, ][n:1,] )
topResults[c(1:5,(2*n-4):(2*n)), c('baseMean','log2FoldChange','padj')]
```

### Plot log fold change vs mean expression for all genes, red = p < 0.05
```{r plotMA}
plotMA(resOrdered, main='DESeq2: BH vs Serum', ylim=c(-2,2), alpha = 0.05)
```

### PCA
```{r plotPCA}
rld <- rlogTransformation(dds, blind=TRUE)
plotPCA(rld)
```

### Effects of transformations on the variance
```{r}
ntd <- normTransform(dds)
meanSdPlot(assay(ntd))
meanSdPlot(assay(rld))

```

### Plot count for the best gene (lowest pval)
```{r plotCounts}
plotCounts(dds, gene=which.min(resOrdered$padj), intgroup='condition', pch = 19 )
```

### Top genes normalized counts heatmap
```{r heatmap}
hmcol <- brewer.pal(11,'RdBu')
nCounts <- counts(dds, normalized=TRUE)
heatmap(as.matrix(nCounts[ row.names(topResults), ]), Rowv = NA, col = hmcol, mar = c(8,2))
```

### Heatmap of the count matrix
```{r}
select <- order(rowMeans(counts(dds,normalized=TRUE)),
                decreasing=TRUE)[1:20]
pheatmap(assay(rld)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE)
pheatmap(assay(ntd)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE)
```

### Sample-to-sample euclidian distance plot
```{r pheatmap}
sampleDists <- dist(t(assay(rld)))
sampleDistMatrix <- as.matrix(sampleDists)
pheatmap(sampleDistMatrix, clustering_distance_rows = sampleDists,
    clustering_distance_cols = sampleDists, color = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255))
```

### ¿How many genes have a padj value smaller than 0.001?
```{r}
filterL2FCG2L05 <- (abs(resOrdered$log2FoldChange) > 2) | (abs(resOrdered$log2FoldChange) < 0.5)
filtered_res <- resOrdered[filterL2FCG2L05,]
sum(filtered_res$padj < 0.001, na.rm=TRUE)
```
